---
title: "Constructing Low-Noise SS3 Datasets for Model Validation"
subtitle: "Expected Data, Low White Noise, and Controlled Misspecification Tests"
author:
  - "L.T. Kell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
    code_folding: hide
    theme: cosmo
    highlight: tango
  word_document:
    toc: true
    toc_depth: 3
    number_sections: true
    reference_docx: reference.docx
editor_options:
  markdown:
    wrap: 72
---

This document provides a reproducible workflow for constructing low-noise SS3 datasets for model validation. The objective is not to create a zero-error dataset, but to create datasets in which observation error is reduced and apparent process-error autocorrelation is minimised, so that deliberately introduced misspecification can be detected more clearly in subsequent analyses. 

The workflow uses the original SS3 fit as the starting point, then creates two derived datasets: a reduced-measurement-error dataset (`meDir`) based on model expected values, and a reduced-measurement plus reduced-process-error dataset (`peDir`) based on catch regularisation. SS3 can output a file of expected observations (`data_expval.ss`) when the starter file requests two data files, and this is the cleanest basis for constructing the low-noise dataset. [web:182]

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  cache      = TRUE,
  cache.path = file.path("cache", paste0(tools::file_path_sans_ext(basename(knitr::current_input())), "/")),
  dev        = "png",
  dpi        = 300,
  fig.width  = 10,
  fig.height = 7
)
```

```{r libs}
library(r4ss)
library(dplyr)
library(ggplot2)
library(plyr)
library(diags)

set.seed(123)

iFig=0
iTab=0

source("C:/active/flr/diags/R/diags-SS.R")
source("C:/active/flr/diags/R/diags-funcs.R")
```

```{r paths}
baseDir="C:/active/flr/sstruth"

ssDir=file.path(baseDir, "data/ss3/herring")
meDir=file.path(baseDir, "data/cond/herring/me")
peDir=file.path(baseDir, "data/cond/herring/pe")

omDir=ssDir
```

## Aim

The purpose is to construct three related model states:

- `ssDir`: the original assessment fit.
- `meDir`: a reduced-measurement-error dataset based on expected values, with optional low white noise added back to observations.
- `peDir`: a reduced-measurement-error dataset with additional catch regularisation intended to reduce apparent autocorrelation in estimated recruitment deviations.

These are synthetic datasets for validation and method comparison, not reconstructions of the true historical system. 

## Workflow

The workflow has five steps:

1. Run the original assessment and request `data_expval.ss` from SS3.
2. Build the reduced-measurement-error dataset in `meDir` from expected values.
3. Fit the reduced-measurement-error model and examine residuals and rec-dev ACF.
4. Build a reduced-process-error dataset in `peDir` by regularising catch.
5. Fit the regularised dataset and compare diagnostics across `ssDir`, `meDir`, and `peDir`. 

## Step 1: Configure SS3 to write expected values

SS3 writes `data_expval.ss` when the starter file requests at least two data files. This file contains expected data values from the fitted model and is the recommended starting point for low-noise synthetic datasets. [web:182]

```{r starter-edit}
starterFile=file.path(omDir, "starter.ss")
strt=readLines(starterFile)
strt=gsub(
  pattern     = "^[[:space:]]*[0-9]+[[:space:]]*# Number of datafiles to produce",
  replacement = " 2 # Number of datafiles to produce",
  x           = strt
)
writeLines(strt, starterFile)

oldwd=getwd()
setwd(omDir)
system("ss3 -nohess", ignore.stdout = TRUE, ignore.stderr = TRUE)
setwd(oldwd)
```

```{r read-expected}
omObs=SS_readdat(file.path(omDir, "data.ss"),        section = 1)
omHat=SS_readdat(file.path(omDir, "data_expval.ss"), section = 1)
```

## Step 2: Build reduced-measurement-error dataset

The `meDir` dataset starts from `data_expval.ss`, because it preserves internal consistency among catch, indices, and composition data while removing the original observational residuals. 

```{r copy-me-files, eval=FALSE}
copy_SS_inputs(
  dir.old    = omDir,
  dir.new    = meDir,
  create.dir = TRUE,
  overwrite  = TRUE,
  use_ss_new = FALSE,
  copy_exe   = TRUE)

meDat=omHat
```

### Optional low white noise

Very small independent noise can be added back to avoid pathological zero-residual behaviour while keeping observation error low and approximately white. This matches the general expected-data plus observation-error construction used in simulation workflows. 

```{r noise-functions}
addLognormalNoise<-function(x, sdlog = 0.02) {
  x * exp(rnorm(length(x), mean = -0.5 * sdlog^2, sd = sdlog))}

rescaleComp<-function(x) {
  s=sum(x, na.rm = TRUE)
  if (is.na(s) || s <= 0) return(x)
  x / s}
```

```{r add-noise}
add_noise  =TRUE
catch_sdlog=0.02
index_sdlog=0.03

if (add_noise && !is.null(meDat$catch)) {
  valid=is.finite(meDat$catch$catch) & meDat$catch$catch > 0
  meDat$catch$catch[valid]=addLognormalNoise(
    meDat$catch$catch[valid],
    sdlog = catch_sdlog
  )
}

if (add_noise && !is.null(meDat$CPUE)) {
  valid=is.finite(meDat$CPUE$obs) & meDat$CPUE$obs > 0
  meDat$CPUE$obs[valid]=addLognormalNoise(
    meDat$CPUE$obs[valid],
    sdlog = index_sdlog
  )
}

# Optional: leave compositions at expected values or add small perturbations here.
```

```{r write-me-dat}
SS_writedat(
  datlist   = meDat,
  outfile   = file.path(meDir, "data.ss"),
  overwrite = TRUE)
```

## Step 3: Fit reduced-measurement-error dataset

```{r run-me, eval=FALSE}
oldwd=getwd()
setwd(meDir)
system("ss3 -nohess", ignore.stdout = TRUE, ignore.stderr = TRUE)
setwd(oldwd)

meFit=SS_output(meDir, covar = TRUE, verbose=FALSE,printstats=FALSE,hidewarn=TRUE)
```

## Step 4: Build reduced-process-error dataset via catch regularisation

The `peDir` dataset starts from `meDir` and applies a catch regularisation step. Here, catch regularisation means smoothing catch in log-space and shrinking the series toward that smooth trend so that apparent autocorrelation in estimated rec devs is reduced. This is a design step for constructing a fit-for-purpose operating-model dataset, not a claim that the resulting catch series is historically correct. 

```{r copy-pe-files, eval=FALSE}
copy_SS_inputs(
  dir.old    = meDir,
  dir.new    = peDir,
  create.dir = TRUE,
  overwrite  = TRUE,
  use_ss_new = FALSE,
  copy_exe   = TRUE)

peDat=SS_readdat(file.path(meDir, "data.ss"), section = 1)
```

```{r catch-regularisation}
if (!is.null(peDat$catch)) {
  x =peDat$catch$catch
  ok=is.finite(x) & x > 0

  lx=log(x[ok])

  k =3L
  w =rep(1 / k, k)
  lx_sm=as.numeric(stats::filter(lx, filter = w, sides = 2))

  na_idx=is.na(lx_sm)
  lx_sm[na_idx]=lx[na_idx]

  alpha =0.5
  lx_adj=(1 - alpha) * lx + alpha * lx_sm

  peDat$catch$catch[ok]=exp(lx_adj)
}

SS_writedat(
  datlist   = peDat,
  outfile   = file.path(peDir, "data.ss"),
  overwrite = TRUE
)
```

```{r run-pe, eval=FALSE}
oldwd=getwd()
setwd(peDir)
system("ss3 -nohess", ignore.stdout = TRUE, ignore.stderr = TRUE)
setwd(oldwd)

peFit=SS_output(peDir, covar = TRUE, verbose=FALSE,printstats=TRUE,hidewarn=TRUE)
```

## Step 5: Diagnostics

Comparing rec-dev ACF across the original fit, reduced-measurement dataset, and reduced-process dataset provides a direct way to assess whether the intended regularisation is actually reducing serial structure in the estimated recruitment deviations. 

```{r diagnostics, eval=FALSE}
ssFit=SS_output(ssDir, covar = TRUE)
meFit=SS_output(meDir, covar = TRUE)
peFit=SS_output(peDir, covar = TRUE)

rec_ss=ssFit$recruit$dev
rec_me=meFit$recruit$dev
rec_pe=peFit$recruit$dev

par(mfrow = c(1, 3))
acf(rec_ss[!is.na(rec_ss)], main = "Original fit")
acf(rec_me[!is.na(rec_me)], main = "Reduced ME")
acf(rec_pe[!is.na(rec_pe)], main = "Reduced ME + PE")
par(mfrow = c(1, 1))
```

## Optional fitted-catch route

If catch is first estimated under an SS3 configuration that treats catch with an error model, the expected catch from `SS_output()$catch$exp` can be written back into the data file before the `meDir` or `peDir` steps. This is an optional alternative route if you want fitted catch rather than `data_expval.ss` catch as the starting series. 

```{r fitted-catch-route, eval=FALSE}
fit=SS_output(ssDir, covar = FALSE)
catch_exp=fit$catch[, c("Yr", "Fleet", "exp")]
names(catch_exp)[names(catch_exp) == "exp"]="exp_catch"

dat=SS_readdat(file.path(ssDir, "data.ss"), section = 1)

for (i in seq_len(nrow(dat$catch))) {
  y =dat$catch$year[i]
  fl=dat$catch$fleet[i]

  idx=which(catch_exp$Yr == y & catch_exp$Fleet == fl)
  if (length(idx) == 1L && is.finite(catch_exp$exp_catch[idx])) {
    dat$catch$catch[i]=catch_exp$exp_catch[idx]
  }
}

SS_writedat(
  datlist   = dat,
  outfile   = file.path(ssDir, "data_catchExp.ss"),
  overwrite = TRUE)
```

## Notes

This workflow is intended to construct an operating-model dataset that is “perfect” for the specific validation task, rather than necessarily “correct” in a historical sense. If rec-dev autocorrelation remains after the `meDir` step, that indicates the remaining structure is not simply observational noise and motivates the additional regularisation step used to construct `peDir`. 
