---
title: "SS Production Error Diagnostics"
subtitle: "PT vs SS Production-Function Comparison"
author:
  - "L.T. Kell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
    code_folding: hide
    theme: cosmo
    highlight: tango
  word_document:
    toc: true
    toc_depth: 3
    number_sections: true
    reference_docx: reference.docx
editor_options:
  markdown:
    wrap: 72
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  cache      = TRUE,
  cache.path = file.path("cache", paste0(tools::file_path_sans_ext(basename(knitr::current_input())), "/")),
  dev        = "png",
  dpi        = 300,
  fig.width  = 10,
  fig.height = 7)
```

## Overview

This workflow compares surplus-production and process-error behavior between:

- SS-derived production-function signals, and
- Pella-Tomlinson fits in biomass and SSB space.

The package now provides generic interfaces:

- `ssPe(x, ...)` for full outputs and plots
- `ssPeDiagnose(x, ...)` for diagnostics only

Where `x` can be:

- an SS3 model directory (`character`), or
- a list with `tseries` and `curve`.

```{r libs}
library(sstruth)
library(r4ss)
```

## Run from an SS3 directory

```{r run-from-dir, eval=FALSE}
ssDir="C:/active/flr/sstruth/data/ss3/herring"

res=ssPe(
  x = ssDir,
  shapePt = 0.2,
  bmsySsb = NULL,
  makePlots = TRUE
)

res$pars
res$diagnostics
```

## Diagnostics-only run

```{r diagnostics-only, eval=FALSE}
diag=ssPeDiagnose(
  x = "C:/active/flr/sstruth/data/ss3/herring",
  shapePt = 0.2
)

diag
```

## Run from prebuilt objects

```{r run-from-list, eval=FALSE}
cs=curveSS("C:/active/flr/sstruth/data/ss3/herring")

res2=ssPe(
  x = list(tseries = cs$tseries, curve = cs$curve),
  shapePt = 0.2,
  makePlots = FALSE
)

res2$diagnostics
```

