---
title: "SS3 to JABBA Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SS3 to JABBA Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, global-options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6
)
```

## Introduction

### What this vignette shows

It walks through one **reproducible path** in `sstruth`: read a fitted Stock Synthesis (SS3) run, translate it to a truth table, build JABBA inputs and (optionally) fit JABBA, then compare SS3 and JABBA on **historical biomass**, **terminal status vs MSY**, **production and yield**, **short-term projections**, and a **simple backtest**. Everything can be driven from `workflow()` plus the `compare*()` and `plot*()` helpers.

### How to interpret the comparisons

**Roles of the two models.** In code, SS3-based quantities use `source = "truth"` (also labelled **SS3** in figures). JABBA outputs use `source = "jabba"` (**JABBA**). SS3 is the **reference** system; JABBA is a simpler biomass model fit to catch and an index from the truth table.

**Colours.** Throughout, **blue** is SS3 and **orange** is JABBA (`cols_ss3_jabba` in the setup chunk).

**Index for JABBA.** `asJabba(..., indexCol = "stock")` uses **exploitable / total biomass** from `asTruth()`, not spawning stock, unless you change `indexCol`.

**SS3 production curve.** When you use `workflow(ssDir, ...)`, `compareProduction()` takes the SS3 side from **`FLRebuild::curveSS(ssDir)`** (equilibrium **yield vs SSB** from the fitted assessment), not a Schaefer regression on the time series. The JABBA side is still a **Schaefer** curve fitted to its catch/biomass series. Call `compareProduction(..., ssDir = NULL)` if you want both sides from the same Schaefer construction.

**Other Schaefer overlays** (e.g. custom vignette chunks) may still fit a **discrete-time Schaefer** to each series for diagnostics; that is separate from the SS3 equilibrium curve above.

**Reading the production and yield figures (Section 7).**

- **Surplus production vs biomass:** **SS3** smooth curve = equilibrium **yield vs SSB** from `curveSS` (not Schaefer). **JABBA** smooth curve = Schaefer `P(B)`. **Path + points** = empirical SP along the observed series (Section 7 uses **SSB** on the x-axis for SS3 so it matches the SS3 curve; **biomass** for JABBA).
- **Dimensionless panels:** **x = B/BMSY** and **y = yield/MSY** (SS3) or **P/MSY** (JABBA). SS3 uses **`interpolate_ss3_curve_scale()`**: **Bden** and **Mden** are **curveSS** refpts when they match the interpolated peak (5% tol), otherwise the **curve peak** — so the equilibrium curve is **(1, 1)** at MSY. **Dotted lines** at **x = 1** and **y = 1**.

**Projections (Section 8).** Biomass is shown as **B/BMSY** using the same Schaefer **BMSY** as the projection dynamics. **Line type** distinguishes **F = 0**, **F = FMSY**, and **F = Fcurrent** (current F from the terminal year, expressed on each model’s scale).

**Runs tests (Section 4a).** JARA residuals are checked for **randomness** (runs and crossings). Shaded facets show whether the usual run-test criteria pass; this is about **index / trend diagnostics**, not SS3 vs JABBA agreement.

### Roadmap

| Section | What you get |
|--------|----------------|
| **1–3** | SS3 directory → `readTruth` / `asTruth` → `asJabba` → `workflow()` |
| **4** | JARA runs tests on diagnostics; `plotAll()` four-panel comparison |
| **5–6** | Historical biomass; terminal-year **B/BMSY** and **F/FMSY** |
| **7** | Production curves, trajectory in production space, yield vs biomass |
| **8** | Forward **B/BMSY** under F0, FMSY, Fcurrent |
| **9** | Retrospective truncation and 1-step-ahead style checks |

**Packages:** `ss3om`, `r4ss`, optional **JABBA** (fit), **JARA** (Section 4a), **FLRebuild** (SS3 equilibrium curve via `curveSS` when using `workflow()`). Set `ssDir` in **Inputs** to your own SS3 run if needed.

### Workflow checklist (i–vi)

| Step | What to use |
|------|-------------|
| **i)** Read SS3 | `readTruth(ssDir)`, `asTruth()` |
| **ii)** JABBA on catch + exploitable biomass | `asJabba(truthDf, indexCol = "stock")`, `fitJabba(..., run = TRUE)` or `workflow(ssDir, runFit = TRUE, indexCol = "stock")` |
| **iii)** SP vs biomass (absolute & relative scaling) | `wf$comparisons$production` (`compareProduction`); Section 7 figures for trajectories and yield vs biomass. SS3 curve from `FLRebuild::curveSS`; JABBA curve = Schaefer on the series. |
| **iv)** As (iii) in **B/BMSY** and **yield/MSY** | Section 7 dimensionless panels: **x = B/BMSY**, **y** scaled by the same **MSY** as **Mden** (SS3) or Schaefer **MSY** (JABBA). |
| **v)** Projections SS3 & JABBA | `wf$comparisons$projections` (`compareProjection`); Section 8 adds Fcurrent via local Schaefer helpers. |
| **vi)** Backtest with **reported catch** | This vignette §9 = simple truncate-and-compare JABBA vs realized series. Full SS3 **reported-catch** forecast scenarios across retros use `runFlevelProjections` / `ss3-retro-projections` tooling, not §9 alone. |

```{r setup}
library(ggplot2)
library(plyr)

library(FLCore)
library(ggplotFL)

library(r4ss)
library(ss3om)
library(sstruth)

# Consistent styling across figures in this vignette.
theme_set(
  theme_bw(base_size = 14) +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "grey95", colour = "grey80"),
      strip.text = element_text(size = 12, face = "bold"),
      axis.title = element_text(size = 13),
      axis.text = element_text(size = 11),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 11),
      plot.title = element_text(face = "bold", size = 16),
      panel.spacing = grid::unit(0.25, "lines")
    )
)

# Colours for SS3 (truth) vs JABBA in comparison figures (Okabe–Ito–style contrast).
cols_ss3_jabba=c("truth" = "#0072B2", "jabba" = "#D55E00")

iFig=0
```

## Inputs

Set your SS3 run directory:

```{r, ssdir-input}
ssDir="C:/active/flr/sstruth/data/ss3/sma-natl/Scenario-1"
```

## 1) Read SS3

```{r, read-ss3, eval=TRUE}
truthObj=readTruth(ssDir)
truthDf =asTruth(truthObj)
```

## 2) Create a JABBA object

JABBA is fed an abundance **index** from the truth table. Here we use **`stock`** (exploitable / total biomass proxy from `asTruth()`), not spawning stock.

```{r, create-jabba-input, eval=TRUE}
jbInp=asJabba(truthDf, indexCol = "stock", cv = 0.2)
```

## 3) Run JABBA

Use the workflow wrapper for a single object containing truth, inputs, fit, and comparisons.

To setup comparisons without fitting JABBA, set `runFit = FALSE`.

```{r, run-workflow, eval=TRUE}
wf=workflow(ssDir, runFit=TRUE, cv=0.2, proj_years=10, indexCol = "stock")
```

## 4) Plot diagnostics

Section 4a builds a JARA object from workflow diagnostics and checks **standardized residuals** for non-random patterns. Section 4b calls **`plotAll(wf$comparisons)`**, which returns **four** ggplots (`historical`, `current_state`, `production`, `projections`). The chunk below **prints three** of them; use `print(plots$current_state)` for the bar chart of terminal **B/BMSY** and **F/FMSY**.

### 4a) Runs tests for the trend analysis

```{r, jara-runs-setup, eval=TRUE}
# Build a real JARA object (hard fail if unavailable).
dgs=diags(wf, "jabba")
dgs=subset(dgs, is.finite(year) & is.finite(obs))
dgs=dgs[order(dgs$name, dgs$year), ]
if (nrow(dgs) == 0) stop("No diagnostics rows available to build JARA input.")
if (!requireNamespace("JARA", quietly = TRUE)) stop("Package 'JARA' is required for this chunk.")

# JARA needs finite observation errors; use provided se where available.
dgs$se_use=ifelse(is.finite(dgs$se) & dgs$se > 0, dgs$se, 0.2)

idx_u=reshape(
  dgs[, c("year", "name", "obs")],
  idvar = "year",
  timevar = "name",
  direction = "wide"
)
idx_se=reshape(
  dgs[, c("year", "name", "se_use")],
  idvar = "year",
  timevar = "name",
  direction = "wide"
)
idx_u=idx_u[order(idx_u$year), ]
idx_se=idx_se[order(idx_se$year), ]

names(idx_u)=sub("^obs\\.", "", names(idx_u))
names(idx_se)=sub("^se_use\\.", "", names(idx_se))
names(idx_u)[1]="Yr"
names(idx_se)[1]="Yr"

sink(file = tempfile())
on.exit(sink(), add = TRUE)
jas=JARA::build_jara(
  I = idx_u,
  se = idx_se,
  variance.weighting = "model",
  model.type = "relative",
  assessment = "workflow",
  GL = 15
)
jas=JARA::fit_jara(jas, do.ppc = TRUE, quickmcmc = TRUE)

# Local run-diagnostics helper matching qic(..., chart = "i")$data fields.
qic=function(year, residual, chart = "i") {
  ord=order(year, na.last = NA)
  rr=as.numeric(residual[ord])
  rr=rr[is.finite(rr)]
  sgn=ifelse(rr >= 0, 1L, -1L)
  n=length(sgn)
  if (n < 2) {
    return(list(data = data.frame(
      n.crossings = NA_real_,
      n.crossings.min = NA_real_,
      longest.run = NA_real_,
      longest.run.max = NA_real_
    )))}
  
  n.crossings=sum(diff(sgn) != 0)
  longest.run=max(rle(sgn)$lengths)
  n.crossings.min=floor((n - 1) * 0.25)
  longest.run.max=max(2, ceiling(log(n, base = 2) + 2))
  list(data = data.frame(
    n.crossings = n.crossings,
    n.crossings.min = n.crossings.min,
    longest.run = longest.run,
    longest.run.max = longest.run.max))}

```

```{r, jara-runs, fig.height=8, eval=TRUE}
rns=ldply(list(jas), function(x) x$fits)
runs=ddply(rns, .(scenario, name), with, qic(year, residual, chart = "i")$data)
runs=ddply(
  runs,
  .(scenario, name),
  with,
  data.frame(
    crossings = all(n.crossings > n.crossings.min),
    runs = all(longest.run < longest.run.max)))

runs=transform(runs, Pass = runs & crossings)

ggplot(rns) +
  geom_hline(aes(yintercept = 0), linewidth = 0.4, colour = "grey40") +
  geom_line(aes(year, residual), position = position_dodge(width = 1), col = "grey50", linewidth = 0.5) +
  geom_point(
    aes(year, residual),
    position = position_dodge(width = 1),
    col = "grey35",
    linewidth = 0.5,
    data = subset(rns, residual < 0)) +
  geom_linerange(
    aes(year, ymin = 0, ymax = residual),
    position = position_dodge(width = 1),
    col = "grey35",
    linewidth = 0.5,
    data = subset(rns, residual < 0)) +
  geom_point(
    aes(year, residual),
    position = position_dodge(width = 1),
    col = "black",
    linewidth = 0.5,
    data = subset(rns, residual > 0)) +
  geom_linerange(
    aes(year, ymin = 0, ymax = residual),
    position = position_dodge(width = 1),
    col = "black",
    linewidth = 0.5,
    data = subset(rns, residual > 0)) +
  facet_grid(name ~ scenario, scale = "free", space = "free_x") +
  theme(
        strip.text.y = element_text(angle = 0, face = "bold"),
        strip.text.x = element_text(angle = 0),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Year") +
  ylab("") +
  geom_rect(
    aes(fill = Pass),
    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf,
    alpha = 0.25,
    data = runs
  ) +
  scale_fill_manual(values = c("FALSE" = "grey80", "TRUE" = "grey35")) +
  scale_x_continuous(breaks = pretty(rns$year, n = 8)) +
  guides(fill = guide_legend(title = "Run test pass"))

```

Figure `r iFig=iFig+1; iFig`. JARA residual runs tests (shading = both runs and crossings criteria satisfied).

### 4b) Comparison plots

```{r, comparison-plots, eval=TRUE}
# Each ggplot must be printed explicitly so knitr records all figures (not only the last).
plots=plotAll(wf$comparisons)
print(plots$historical)
print(plots$production)
print(plots$projections)
```

## 5) Compare historical state

Absolute biomass from each model (same units as in the truth table and JABBA output), not scaled to MSY.

```{r, compare-historical, eval=TRUE}
hist=wf$comparisons$historical
ggplot(hist)+
  geom_line(aes(year, truth_stock, color = "SS3"), linewidth = 0.9)+
  geom_line(aes(year, jabba_stock, color = "JABBA"), linewidth = 0.9)+
  scale_color_manual(
    values = c("SS3" = cols_ss3_jabba["truth"], "JABBA" = cols_ss3_jabba["jabba"]),
    name = NULL
  ) +
  labs(x = "Year", y = "Biomass", title = "Historical biomass: SS3 vs JABBA")
```

## 6) Current state relative to reference points

Terminal year of the merged history. **B/BMSY** and **F/FMSY** here use **Schaefer** `BMSY` and `FMSY` from each **time series** (`compareCurrent()`), not SS3 report reference points. That can differ from **`curveSS` refpts** used on the SS3 production curve in `compareProduction()`.

```{r, current-state-table, eval=TRUE}
cur=wf$comparisons$current_state

```

Columns in `current_state`:

- **`truth_B_Bmsy`**, **`jabba_B_Bmsy`** — biomass relative to that series’ Schaefer **BMSY** (0.5·**K**).
- **`truth_F_Fmsy`**, **`jabba_F_Fmsy`** — **catch/biomass** in the terminal year relative to that series’ **FMSY** (0.5·**r**).

Values near **1** mean “at the Schaefer MSY reference” for that quantity; they are **not** SS3’s reported BMSY/FMSY unless the Schaefer approximation matches the assessment.

## 7) Production functions with stock trajectory

Object `wf$comparisons$production` combines **`source = "truth"`** rows from **`FLRebuild::curveSS(ssDir)`** (equilibrium **yield vs SSB**; `r` and `K` are `NA`) with **`source = "jabba"`** rows from a **Schaefer** grid (`stock`, `production`, `r`, `K`, …). A quick look:

```{r, production-table, eval=TRUE}
prod=wf$comparisons$production
head(prod)
```

Absolute-scale production functions: **x = B/BMSY** for both series (SS3 uses `curveSS` refpts or interpolated peak denominator, same as relative panels; JABBA uses Schaefer **BMSY**). **y** remains yield or empirical surplus production in model units. `ss_ref` is MSY/BMSY from `curveSS` when `prod` has `r`/`K` = `NA` for SS3.

```{r, production-with-trajectory, eval=TRUE}
hist=wf$comparisons$historical

cs_ss=if (requireNamespace("FLRebuild", quietly = TRUE)) {
  if (!"package:dplyr" %in% search()) suppressPackageStartupMessages(library(dplyr))
  FLRebuild::curveSS(ssDir)
} else NULL

rp=if (!is.null(cs_ss) && is.data.frame(cs_ss$refpts) && nrow(cs_ss$refpts) > 0) cs_ss$refpts else NULL
ss_ref=list(
  Bmsy = if (!is.null(rp)) as.numeric(rp$bmsy[1]) else NA_real_,
  MSY = if (!is.null(rp)) as.numeric(rp$msy[1]) else NA_real_
)

bmsy_schaefer=function(biomass, catch) {
  ok=is.finite(biomass) & is.finite(catch) & biomass > 0
  biomass=biomass[ok]
  catch=catch[ok]
  fit=lm(
    I(
      ((biomass[-1] - biomass[-length(biomass)]) + catch[-length(catch)]) /
        pmax(biomass[-length(biomass)], 1e-12)
    ) ~ biomass[-length(biomass)]
  )
  r=coef(fit)[1]
  slope=coef(fit)[2]
  K=-r / slope
  0.5 * as.numeric(K)
}

bmsy_lookup=data.frame(
  source = c("truth", "jabba"),
  Bmsy = c(
    if (is.finite(ss_ref$Bmsy)) ss_ref$Bmsy else bmsy_schaefer(hist$truth_stock, hist$truth_catch),
    bmsy_schaefer(hist$jabba_stock, hist$jabba_catch)
  ),
  stringsAsFactors = FALSE
)

# Interpolate SS3 curve; Bden on x is refpts BMSY if that matches the peak (tol), else curve peak.
interpolate_ss3_curve_scale=function(cs_ss, ss_ref, tol = 0.05) {
  cv=cs_ss$curve
  cv=cv[is.finite(cv$ssb) & is.finite(cv$yield) & cv$ssb >= 0, ]
  if (nrow(cv) < 2L) return(NULL)
  cv=stats::aggregate(stats::as.formula("yield ~ ssb"), cv, function(z) mean(z, na.rm = TRUE))
  cv=cv[order(cv$ssb), , drop = FALSE]
  ssg=seq(min(cv$ssb), max(cv$ssb), length.out = 500)
  yg=stats::approx(cv$ssb, cv$yield, xout = ssg)$y
  im=which.max(yg)
  Bpeak=ssg[im]
  Mpeak=yg[im]
  Bref=ss_ref$Bmsy
  Mref=ss_ref$MSY
  ref_ok=is.finite(Bref) && is.finite(Mref) && Bref > 0 && Mref > 0
  use_ref=ref_ok && abs(Bpeak - Bref) / Bref < tol && abs(Mpeak - Mref) / Mref < tol
  Bden=if (use_ref) Bref else Bpeak
  Mden=if (use_ref) Mref else Mpeak
  list(
    ssg = ssg, yg = yg, Bden = Bden, Mden = Mden, use_ref = use_ref,
    Bref = if (ref_ok) Bref else NA_real_, Mref = if (ref_ok) Mref else NA_real_,
    Bpeak = Bpeak, Mpeak = Mpeak
  )
}

ss3_curve_scale=if (!is.null(cs_ss)) interpolate_ss3_curve_scale(cs_ss, ss_ref) else NULL

Bden_truth=if (!is.null(ss3_curve_scale)) ss3_curve_scale$Bden else bmsy_lookup$Bmsy[bmsy_lookup$source == "truth"]
Mden_truth=if (!is.null(ss3_curve_scale)) {
  ss3_curve_scale$Mden
} else if (is.finite(ss_ref$MSY)) {
  ss_ref$MSY
} else {
  pt0=prod[prod$source == "truth", , drop = FALSE]
  if (nrow(pt0) && any(is.finite(pt0$production))) max(pt0$production, na.rm = TRUE) else NA_real_
}
Bmsy_j_abs=bmsy_lookup$Bmsy[bmsy_lookup$source == "jabba"]
pj=prod[prod$source == "jabba", , drop = FALSE]
rj=if (nrow(pj) && any(is.finite(pj$r))) stats::na.omit(pj$r)[1] else NA_real_
Kj=if (nrow(pj) && any(is.finite(pj$K))) stats::na.omit(pj$K)[1] else NA_real_
MSY_j=if (is.finite(rj) && is.finite(Kj)) rj * Kj / 4 else if (nrow(pj) && any(is.finite(pj$production))) max(pj$production, na.rm = TRUE) else NA_real_

prod_plot=rbind(
  transform(prod[prod$source == "truth", , drop = FALSE], x_plot = stock / Bden_truth),
  transform(prod[prod$source == "jabba", , drop = FALSE], x_plot = stock / Bmsy_j_abs)
)

# Dimensionless: SS3 uses interpolated ssg/yg so the equilibrium curve peaks at (1, 1); JABBA uses prod grid.
prod_truth_dim=if (!is.null(ss3_curve_scale)) {
  sc=ss3_curve_scale
  data.frame(source = "truth", x_plot = sc$ssg / sc$Bden, y_plot = sc$yg / sc$Mden)
} else {
  pt=prod[prod$source == "truth", , drop = FALSE]
  data.frame(source = "truth", x_plot = pt$stock / Bden_truth, y_plot = pt$production / Mden_truth)
}
prod_plot_dim=rbind(
  prod_truth_dim,
  data.frame(source = "jabba", x_plot = pj$stock / Bmsy_j_abs, y_plot = pj$production / MSY_j)
)

traj_truth=transform(
  hist,
  source = "truth",
  stock = truth_ssb,
  production = c(diff(truth_ssb) + head(truth_catch, -1), NA)
)[, c("year", "source", "stock", "production")]

traj_jabba=transform(
  hist,
  source = "jabba",
  stock = jabba_stock,
  production = c(diff(jabba_stock) + head(jabba_catch, -1), NA)
)[, c("year", "source", "stock", "production")]

traj=rbind(traj_truth, traj_jabba)
traj=traj[is.finite(traj$stock) & is.finite(traj$production), ]

traj_plot=rbind(
  transform(traj[traj$source == "truth", , drop = FALSE], x_plot = stock / Bden_truth),
  transform(traj[traj$source == "jabba", , drop = FALSE], x_plot = stock / Bmsy_j_abs)
)

traj_plot_dim=rbind(
  transform(traj[traj$source == "truth", , drop = FALSE], x_plot = stock / Bden_truth, y_plot = production / Mden_truth),
  transform(traj[traj$source == "jabba", , drop = FALSE], x_plot = stock / Bmsy_j_abs, y_plot = production / MSY_j)
)

ggplot() +
  geom_line(
    data = prod_plot,
    aes(x = x_plot, y = production, color = source),
    linewidth = 1.1
  ) +
  geom_path(
    data = traj_plot,
    aes(x = x_plot, y = production, color = source),
    alpha = 0.85,
    linewidth = 0.8
  ) +
  geom_point(
    data = traj_plot,
    aes(x = x_plot, y = production, color = source),
    size = 1.6,
    alpha = 0.9
  ) +
  scale_color_manual(
    values = cols_ss3_jabba,
    labels = c("truth" = "SS3", "jabba" = "JABBA"),
    name = NULL
  ) +
  labs(
    title = "Production functions with stock trajectory (absolute scale)",
    subtitle = "x = B/BMSY: SS3 uses curveSS refpts or interpolated peak (same BMSY as relative panels); JABBA uses Schaefer BMSY.",
    x = "B/BMSY",
    y = "Yield or surplus production",
    color = "Source"
  )
```

**Dimensionless:** same **Bden** / **Mden** as `interpolate_ss3_curve_scale()` so the SS3 equilibrium curve peaks at **(1, 1)**; trajectories use **yield/MSY** and **P/MSY** on **y**. Uses **`prod_plot_dim`** / **`traj_plot_dim`**.

```{r, production-with-trajectory-relative, eval=TRUE}
sc=ss3_curve_scale
subt=if (is.null(sc)) {
  "SS3: no curveSS grid; Bden/Mden from refpts or curve peak / max yield."
} else if (isTRUE(sc$use_ref)) {
  "SS3: B/BMSY and yield/MSY use curveSS refpts (match interpolated peak)."
} else {
  sprintf("SS3: B/BMSY and yield/MSY use curve peak (B=%.4g, MSY=%.4g); refpts BMSY=%.4g.", sc$Bpeak, sc$Mpeak, sc$Bref)
}

ggplot() +
  geom_vline(xintercept = 1, linewidth = 0.5, linetype = "dotted", colour = "grey35") +
  geom_hline(yintercept = 1, linewidth = 0.5, linetype = "dotted", colour = "grey35") +
  geom_line(data = prod_plot_dim, aes(x = x_plot, y = y_plot, color = source), linewidth = 1.1) +
  geom_path(data = traj_plot_dim, aes(x = x_plot, y = y_plot, color = source), alpha = 0.85, linewidth = 0.8) +
  geom_point(data = traj_plot_dim, aes(x = x_plot, y = y_plot, color = source), size = 1.6, alpha = 0.9) +
  scale_color_manual(
    values = cols_ss3_jabba,
    labels = c("truth" = "SS3", "jabba" = "JABBA"),
    name = NULL
  ) +
  labs(
    title = "Production: yield/MSY or P/MSY vs B/BMSY",
    subtitle = subt,
    x = "B/BMSY",
    y = "Yield/MSY or P/MSY",
    color = "Source"
  )
```

Annual **catch** (yield) vs biomass with the same Schaefer **surplus-production** curve `P(B)` for reference (absolute scale):

```{r, yield-vs-biomass-absolute, eval=TRUE}
msy_df=unique(prod[, c("source", "r", "K")])
msy_df$MSY=msy_df$r * msy_df$K / 4
if (is.finite(ss_ref$MSY)) {
  msy_df$MSY[msy_df$source == "truth" & !is.finite(msy_df$MSY)]=ss_ref$MSY
}

yield_traj=rbind(
  transform(hist, source = "truth", biomass = truth_ssb, yield = truth_catch),
  transform(hist, source = "jabba", biomass = jabba_stock, yield = jabba_catch)
)
yield_traj=yield_traj[is.finite(yield_traj$biomass) & is.finite(yield_traj$yield), ]

ggplot() +
  geom_line(
    data = prod,
    aes(x = stock, y = production, color = source),
    linewidth = 1.05,
    linetype = "solid"
  ) +
  geom_path(
    data = yield_traj,
    aes(x = biomass, y = yield, color = source),
    linewidth = 0.85,
    linetype = "dashed"
  ) +
  geom_point(
    data = yield_traj,
    aes(x = biomass, y = yield, color = source),
    size = 1.4
  ) +
  scale_color_manual(
    values = cols_ss3_jabba,
    labels = c("truth" = "SS3", "jabba" = "JABBA"),
    name = NULL
  ) +
  labs(
    title = "Yield (annual catch) vs SSB/biomass with production curves (absolute scale)",
    subtitle = "Solid: SS3 equilibrium yield (curveSS) or JABBA Schaefer P(B); dashed + points: observed catch",
    x = "SSB (SS3) or biomass (JABBA)",
    y = "Mass per year",
    color = "Source"
  )
```

The same comparison in **dimensionless** form: **catch/MSY** and **yield/MSY** on **y**, with **`prod_plot_dim`** for the solid equilibrium curves.

```{r, yield-vs-biomass-relative, eval=TRUE}
prod_y=data.frame(source = prod_plot_dim$source, x_rel = prod_plot_dim$x_plot, yP_rel = prod_plot_dim$y_plot)
yield_rel=rbind(
  data.frame(source = "truth", x_rel = hist$truth_ssb / Bden_truth, y_rel = hist$truth_catch / Mden_truth),
  data.frame(source = "jabba", x_rel = hist$jabba_stock / Bmsy_j_abs, y_rel = hist$jabba_catch / MSY_j)
)
yield_rel=yield_rel[is.finite(yield_rel$x_rel) & is.finite(yield_rel$y_rel), ]

ggplot() +
  geom_vline(xintercept = 1, linewidth = 0.5, linetype = "dotted", colour = "grey35") +
  geom_hline(yintercept = 1, linewidth = 0.5, linetype = "dotted", colour = "grey35") +
  geom_line(data = prod_y, aes(x = x_rel, y = yP_rel, color = source), linewidth = 1.05, linetype = "solid") +
  geom_path(data = yield_rel, aes(x = x_rel, y = y_rel, color = source), linewidth = 0.85, linetype = "dashed") +
  geom_point(data = yield_rel, aes(x = x_rel, y = y_rel, color = source), size = 1.4) +
  scale_color_manual(
    values = cols_ss3_jabba,
    labels = c("truth" = "SS3", "jabba" = "JABBA"),
    name = NULL
  ) +
  labs(
    title = "Yield vs B/BMSY: catch/MSY and yield/MSY",
    subtitle = "Solid: equilibrium curve; dashed + points: observed catch. Same Bden/Mden as production dimensionless panel.",
    x = "B/BMSY",
    y = "Catch/MSY or yield/MSY",
    color = "Source"
  )
```

## 8) Projection scenarios: FMSY, Fcurrent, F0

**Built-in** `compareProjection()` supplies **F0** and **FFMSY** only. Below, **`Fcurrent`** is added with local helpers using the **same Schaefer** parameters as elsewhere. **Facet** = model; **line type** = fishing scenario; horizontal dotted line = **B/BMSY = 1**.

```{r, projection-setup-and-run, eval=TRUE}
fit_schaefer_local=function(year, biomass, catch) {
  ok=is.finite(year) & is.finite(biomass) & is.finite(catch) & biomass > 0
  year=year[ok]; biomass=biomass[ok]; catch=catch[ok]
  ord=order(year)
  biomass=biomass[ord]; catch=catch[ord]
  fit=lm(
    I(
      ((biomass[-1] - biomass[-length(biomass)]) + catch[-length(catch)]) /
        pmax(biomass[-length(biomass)], 1e-12)
    ) ~ biomass[-length(biomass)]
  )
  r=coef(fit)[1]
  slope=coef(fit)[2]
  K=-r / slope
  list(r = as.numeric(r), K = as.numeric(K), Bmsy = 0.5 * as.numeric(K), Fmsy = 0.5 * as.numeric(r))
}

project_local=function(year0, B0, pars, nyears, Fval) {
  yrs=seq.int(year0 + 1, by = 1, length.out = nyears)
  B=numeric(nyears)
  C=numeric(nyears)
  Bt=B0
  for (i in seq_len(nyears)) {
    Ct=Fval * Bt
    Bt=max(Bt + pars$r * Bt * (1 - Bt / pars$K) - Ct, 1e-10)
    B[i]=Bt
    C[i]=Ct
  }
  data.frame(year = yrs, stock = B, catch = C, F = Fval)
}

hist=wf$comparisons$historical
cur =wf$comparisons$current_state
yterm   =max(hist$year, na.rm = TRUE)

t_pars=fit_schaefer_local(hist$year, hist$truth_stock, hist$truth_catch)
j_pars=fit_schaefer_local(hist$year, hist$jabba_stock, hist$jabba_catch)

Fcur_truth=with(cur, truth_F_Fmsy * t_pars$Fmsy)
Fcur_jabba=with(cur, jabba_F_Fmsy * j_pars$Fmsy)

proj_truth=rbind(
  transform(project_local(yterm, cur$truth_stock, t_pars, nyears = 10, Fval = 0), scenario = "F0", source = "truth"),
  transform(project_local(yterm, cur$truth_stock, t_pars, nyears = 10, Fval = t_pars$Fmsy), scenario = "FFMSY", source = "truth"),
  transform(project_local(yterm, cur$truth_stock, t_pars, nyears = 10, Fval = Fcur_truth), scenario = "Fcurrent", source = "truth")
)

proj_jabba=rbind(
  transform(project_local(yterm, cur$jabba_stock, j_pars, nyears = 10, Fval = 0), scenario = "F0", source = "jabba"),
  transform(project_local(yterm, cur$jabba_stock, j_pars, nyears = 10, Fval = j_pars$Fmsy), scenario = "FFMSY", source = "jabba"),
  transform(project_local(yterm, cur$jabba_stock, j_pars, nyears = 10, Fval = Fcur_jabba), scenario = "Fcurrent", source = "jabba")
)

proj_all=rbind(proj_truth, proj_jabba)
proj_all$Bmsy=ifelse(proj_all$source == "truth", t_pars$Bmsy, j_pars$Bmsy)
proj_all$stock_over_Bmsy=ifelse(
  is.finite(proj_all$Bmsy) & proj_all$Bmsy > 0,
  proj_all$stock / proj_all$Bmsy,
  NA_real_
)
proj_all
```

Plot:

```{r, projection-plot, eval=TRUE}
ggplot(proj_all, aes(year, stock_over_Bmsy)) +
  geom_line(aes(linetype = scenario, color = source), linewidth = 1.0) +
  geom_point(aes(shape = scenario, color = source), size = 1.9, stroke = 0.6) +
  geom_hline(yintercept = 1, linewidth = 0.5, linetype = "dotted", colour = "grey35") +
  facet_grid(~source, labeller = ggplot2::labeller(source = c(truth = "SS3", jabba = "JABBA"))) +
  scale_color_manual(values = cols_ss3_jabba, guide = "none") +
  scale_linetype_manual(values = c("F0" = "solid", "Fcurrent" = "dashed", "FFMSY" = "dotdash")) +
  scale_shape_manual(values = c("F0" = 16, "Fcurrent" = 17, "FFMSY" = 15)) +
  labs(
    title = "Projections under F0, Fcurrent, and FFMSY relative to BMSY",
    x = "Year",
    y = "B/BMSY",
    linetype = "Scenario",
    shape = "Scenario"
  )
```

## 9) Backtest

For each **terminal year**, the truth table is cut at that year, JABBA is set up on the truncated series (`runFit = FALSE` here), and terminal **stock** from SS3 vs JABBA is compared to the **realized** stock in the following year. This is a **minimal** retrospective check, not a full MSE.

```{r, backtest-run, eval=TRUE}
backtest_years=function(ssDir, terminal_years, runFit = FALSE, cv = 0.2) {
  out=vector("list", length(terminal_years))
  names(out)=terminal_years

  base_truth=sstruth::asTruth(sstruth::readTruth(ssDir))

  for (i in seq_along(terminal_years)) {
    y_end=terminal_years[i]
    trn=base_truth[base_truth$year <= y_end, , drop = FALSE]
    tst=base_truth[base_truth$year == (y_end + 1), , drop = FALSE]
    if (nrow(tst) == 0) next

    jin=sstruth::asJabba(trn, indexCol = "stock", cv = cv)
    jfit=sstruth::fitJabba(jin, run = runFit)

    cur=compareCurrent(trn, jfit, jin)

    out[[i]]=data.frame(
      terminal_year = y_end,
      realized_year = y_end + 1,
      realized_stock = tst$stock[1],
      realized_catch = tst$catch[1],
      truth_terminal_stock = cur$truth_stock[1],
      jabba_terminal_stock = cur$jabba_stock[1]
    )
  }
  do.call(rbind, out)
}

bt=backtest_years(ssDir, terminal_years = 2005:2015, runFit = FALSE)
bt
```

---

**Using pieces on their own:** `readTruth`, `asTruth`, `asJabba`, `fitJabba`, `diags`, `compare*`, and `plot*` each work independently; `workflow()` bundles truth, inputs, optional fit, and `compareAll()` in one list. SS3 **production** in `compareProduction()` comes from **`FLRebuild::curveSS()`** when `ssDir` is set (as in `workflow()`); JABBA’s curve is still **Schaefer** from the series. Custom plots below may use additional Schaefer overlays; the **index** choice (`stock` here) and **`runFit`** still matter for JABBA.
