---
title: "SS3 to JABBA Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SS3 to JABBA Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

This vignette shows the intended end-to-end workflow in `sstruth`:

1. read in SS3
2. create a JABBA object
3. run JABBA
4. plot diagnostics
5. compare historical state
6. compare current state relative to reference points
7. plot production functions with stock trajectory
8. run projections for `F = FMSY`, `F = Fcurrent`, and `F = 0`
9. run a simple backtest

```{r setup}
library(ggplot2)

library(FLCore)
library(ggplotFL)

library(r4ss)
library(ss3om)
library(sstruth)
```

## Inputs

Set your SS3 run directory:

```{r}
ssDir="C:/active/flr/sstruth/data/ss3/sma-natl/Scenario-1"
```

## 1) Read SS3

```{r, eval=TRUE}
truthObj=readTruth(ssDir)
truthDf =asTruth(truthObj)
head(truthDf)
```

## 2) Create a JABBA object

```{r, eval=TRUE}
jbInp=asJabba(truthDf, indexCol = "ssb", cv = 0.2)
str(jbInp)
```

## 3) Run JABBA

Use the workflow wrapper for a single object containing truth, inputs, fit, and comparisons.

```{r, eval=TRUE}
wf=workflow(ssDir, runFit = TRUE, cv = 0.2, proj_years = 10)
```

If you only want setup/comparisons without fitting JABBA, set `runFit = FALSE`.

## 4) Plot diagnostics

### 4a) Standardized residual diagnostics table

```{r, eval=TRUE}
dgs=rbind(diags(ssDir, "ss"),
          diags(wf, "jabba"))

ggplot(dgs)+
  geom_point(aes(year,residual,col=method))+
  facet_wrap(~name,scale="free")

```

### 4b) Comparison plots

```{r, eval=TRUE}
plots=plotAll(wf$comparisons)
plots$historical
plots$current_state
plots$production
plots$projections
```

## 5) Compare historical state

```{r, eval=TRUE}
hist=wf$comparisons$historical
ggplot(hist)+
  geom_line(aes(year,truth_stock))+
  geom_line(aes(year,jabba_stock))
```

## 6) Current state relative to reference points

```{r, eval=TRUE}
cur=wf$comparisons$current_state
cur
```

The `current_state` table includes:

- `truth_B_Bmsy`, `jabba_B_Bmsy`
- `truth_F_Fmsy`, `jabba_F_Fmsy`

## 7) Production functions with stock trajectory

The production-function comparison table is already available:

```{r, eval=TRUE}
prod=wf$comparisons$production
head(prod)
```

To overlay the observed stock trajectory on top of the production curves:

```{r, eval=TRUE}
hist=wf$comparisons$historical

traj_truth=transform(
  hist,
  source = "truth",
  stock = truth_stock,
  production = c(diff(truth_stock) + head(truth_catch, -1), NA)
)[, c("year", "source", "stock", "production")]

traj_jabba=transform(
  hist,
  source = "jabba",
  stock = jabba_stock,
  production = c(diff(jabba_stock) + head(jabba_catch, -1), NA)
)[, c("year", "source", "stock", "production")]

traj=rbind(traj_truth, traj_jabba)
traj=traj[is.finite(traj$stock) & is.finite(traj$production), ]

ggplot() +
  geom_line(
    data = prod,
    aes(x = stock, y = production, color = source),
    linewidth = 1
  ) +
  geom_path(
    data = traj,
    aes(x = stock, y = production, color = source),
    alpha = 0.6
  ) +
  geom_point(
    data = traj,
    aes(x = stock, y = production, color = source),
    size = 1
  ) +
  theme_minimal() +
  labs(
    title = "Production functions with stock trajectory",
    x = "Biomass",
    y = "Surplus production",
    color = NULL
  )
```

## 8) Projection scenarios: FMSY, Fcurrent, F0

`compareProjection()` currently returns `F0` and `FFMSY`.  
This section adds `Fcurrent` with a local helper that uses the same Schaefer form.

```{r, eval=TRUE}
fit_schaefer_local=function(year, biomass, catch) {
  ok=is.finite(year) & is.finite(biomass) & is.finite(catch) & biomass > 0
  year=year[ok]; biomass=biomass[ok]; catch=catch[ok]
  ord=order(year)
  biomass=biomass[ord]; catch=catch[ord]
  fit=lm(
    I(
      ((biomass[-1] - biomass[-length(biomass)]) + catch[-length(catch)]) /
        pmax(biomass[-length(biomass)], 1e-12)
    ) ~ biomass[-length(biomass)]
  )
  r=coef(fit)[1]
  slope=coef(fit)[2]
  K=-r / slope
  list(r = as.numeric(r), K = as.numeric(K), Bmsy = 0.5 * as.numeric(K), Fmsy = 0.5 * as.numeric(r))
}

project_local=function(year0, B0, pars, nyears, Fval) {
  yrs=seq.int(year0 + 1, by = 1, length.out = nyears)
  B=numeric(nyears)
  C=numeric(nyears)
  Bt=B0
  for (i in seq_len(nyears)) {
    Ct=Fval * Bt
    Bt=max(Bt + pars$r * Bt * (1 - Bt / pars$K) - Ct, 1e-10)
    B[i]=Bt
    C[i]=Ct
  }
  data.frame(year = yrs, stock = B, catch = C, F = Fval)
}

hist=wf$comparisons$historical
cur =wf$comparisons$current_state
yterm   =max(hist$year, na.rm = TRUE)

t_pars=fit_schaefer_local(hist$year, hist$truth_stock, hist$truth_catch)
j_pars=fit_schaefer_local(hist$year, hist$jabba_stock, hist$jabba_catch)

Fcur_truth=with(cur, truth_F_Fmsy * t_pars$Fmsy)
Fcur_jabba=with(cur, jabba_F_Fmsy * j_pars$Fmsy)

proj_truth=rbind(
  transform(project_local(yterm, cur$truth_stock, t_pars, nyears = 10, Fval = 0), scenario = "F0", source = "truth"),
  transform(project_local(yterm, cur$truth_stock, t_pars, nyears = 10, Fval = t_pars$Fmsy), scenario = "FFMSY", source = "truth"),
  transform(project_local(yterm, cur$truth_stock, t_pars, nyears = 10, Fval = Fcur_truth), scenario = "Fcurrent", source = "truth")
)

proj_jabba=rbind(
  transform(project_local(yterm, cur$jabba_stock, j_pars, nyears = 10, Fval = 0), scenario = "F0", source = "jabba"),
  transform(project_local(yterm, cur$jabba_stock, j_pars, nyears = 10, Fval = j_pars$Fmsy), scenario = "FFMSY", source = "jabba"),
  transform(project_local(yterm, cur$jabba_stock, j_pars, nyears = 10, Fval = Fcur_jabba), scenario = "Fcurrent", source = "jabba")
)

proj_all=rbind(proj_truth, proj_jabba)
proj_all
```

Plot:

```{r, eval=TRUE}
ggplot(proj_all, aes(year, stock, color = scenario)) +
  geom_line(linewidth = 1) +
  facet_wrap(~source, scales = "free_y") +
  theme_minimal() +
  labs(title = "Projections under F0, Fcurrent, FFMSY", y = "Biomass", color = NULL)
```

## 9) Backtest

A simple retrospective-style backtest truncates the terminal year, reruns the workflow, and compares 1-step-ahead predictions against realized values.

```{r, eval=TRUE}
backtest_years=function(ssDir, terminal_years, runFit = FALSE, cv = 0.2) {
  out=vector("list", length(terminal_years))
  names(out)=terminal_years

  base_truth=sstruth::asTruth(sstruth::readTruth(ssDir))

  for (i in seq_along(terminal_years)) {
    y_end=terminal_years[i]
    trn=base_truth[base_truth$year <= y_end, , drop = FALSE]
    tst=base_truth[base_truth$year == (y_end + 1), , drop = FALSE]
    if (nrow(tst) == 0) next

    jin=sstruth::asJabba(trn, indexCol = "ssb", cv = cv)
    jfit=sstruth::fitJabba(jin, run = runFit)

    cur=compareCurrent(trn, jfit, jin)

    out[[i]]=data.frame(
      terminal_year = y_end,
      realized_year = y_end + 1,
      realized_stock = tst$stock[1],
      realized_catch = tst$catch[1],
      truth_terminal_stock = cur$truth_stock[1],
      jabba_terminal_stock = cur$jabba_stock[1]
    )
  }
  do.call(rbind, out)
}

bt=backtest_years(ssDir, terminal_years = 2005:2015, runFit = FALSE)
bt
```

---

This workflow is designed so each component can be used standalone (`readTruth`, `asJabba`, `fitJabba`, `diags`, `compare*`, `plot*`) or through `workflow()` for a single end-to-end object.
